---
title: "p8105_hw6_kb3066"
author: "Kaila Boyer"
date: "11/28/2021"
output: github_document
---

```{r setup, include=FALSE}

library(tidyverse)
library(modelr) 
library(mgcv)


knitr::opts_chunk$set(
  fig.width = 7,
  fig.asp = .7,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis",
  scale_colour_discrete = scale_colour_viridis_d,
scale_fill_discrete = scale_fill_viridis_d
)
```
## Problem 1 

```{r}
birth <- read_csv("birthweight.csv") %>% 
                            janitor::clean_names() %>% 
                            mutate(
                              babysex = as.factor(babysex),
                              babysex = fct_recode(babysex, "Male" = "1", "Female" = "2"),
                              frace = as.factor(frace),
                              frace = fct_recode(frace, "White" = "1", "Black" = "2", "Asian" = "3", 
                                                 "Puerto Rican" = "4", "Other" = "8"),
                              malform = as.logical(malform),
                              mrace = as.factor(mrace),
                              mrace = fct_recode(mrace, "White" = "1", "Black" = "2", "Asian" = "3", 
                                                 "Puerto Rican" = "4"))


```


```{r}
fit1 <- lm(bwt ~ blength + babysex,  data = birth )
fit1 %>% broom::tidy()

modelr::add_residuals(birth, fit1) %>% 
  ggplot(aes(y= resid, x = babysex)) +
  geom_point() 

modelr::add_predictions(birth, fit1) %>% 
  ggplot(aes(x = babysex, y = pred)) + 
  geom_point() 
```
The model I proposed was that length at birth and sex of the child help predict birth weight. Length of an object, particularly an infant, would likely have a direct relationship with length (a longer baby would weight more). Sex-differences can be observed with health outcomes, so I included sex of the baby. 

```{r}
fit2 <- lm(bwt ~ blength + gaweeks, data = birth)
fit2 %>% broom::tidy()

fit3 <- lm(bwt ~ bhead + blength + babysex + bhead * blength
           + bhead * babysex + blength * babysex + bhead * babysex * blength, data = birth)
fit3 %>% broom::tidy() %>% knitr::kable(digits = 3)
```

Visual comparison of models (residuals and predictions)
```{r}
birth %>% 
  gather_predictions(fit1, fit2, fit3) %>% 
  mutate(model = fct_inorder(model)) %>%  
  ggplot(aes(x = bwt, y = pred)) + 
  geom_point(alpha = .5) +
  geom_smooth( color = "red") + 
  facet_grid(~model)

birth %>% 
  gather_residuals(fit1, fit2, fit3) %>% 
  mutate(model = fct_inorder(model)) %>%  
  ggplot(aes(x = bwt, y = resid)) + 
  geom_point(alpha = .5) +
  geom_smooth( color = "red") + 
  facet_grid(~model)
```

Comparison and cross-validation of models
```{r}
cv_df <- crossv_mc(birth, 1000) %>% 
  mutate(
    train = map(train, as_tibble), 
    test = map(test, as_tibble)
  )

cv_df = 
  cv_df %>% 
  mutate(
    fit1_mod = map(train, ~lm(bwt ~ blength + smoken + mrace, data = .x)),
    fit2_mod = map(train, ~lm(bwt ~ blength + gaweeks, data = .x)),
    fit3_mod = map(train, ~lm(bwt ~ bhead + blength + babysex + bhead * blength + bhead * babysex + 
                              blength * babysex + bhead * babysex * blength, data = .x))) %>% 
  mutate(
    rmse_fit1 = map2_dbl(fit1_mod, test, ~rmse(model = .x, data = .y)),
    rmse_fit2 = map2_dbl(fit2_mod, test, ~rmse(model = .x, data = .y)),
    rmse_fit3 = map2_dbl(fit3_mod, test, ~rmse(model = .x, data = .y)))

cv_df %>% 
  select(starts_with("rmse")) %>% 
  pivot_longer(
    everything(),
    names_to = "model", 
    values_to = "rmse",
    names_prefix = "rmse_") %>% 
  mutate(model = fct_inorder(model)) %>% 
  ggplot(aes(x = model, y = rmse)) + geom_violin()
```

## Problem 2

```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
```

Bootstrap estimates
```{r}
boot_sample = function(df) {
  sample_frac(df, replace = TRUE)
}

weather_bootstraps = 
  data_frame(
    strap_number = 1:5000,
    strap_sample = rerun(5000, boot_sample(weather_df)))

results <-weather_bootstraps %>%  
                    mutate(
                        models = map(strap_sample, ~lm(tmax ~ tmin, data = .x) ),
                        results = map(models, broom::tidy)) %>% 
                    select(-strap_sample, -models) %>% 
                    unnest(results) 

head(results) %>% knitr::kable()

results %>% 
  group_by(term) %>% 
  summarize(
    ci_lower = quantile(estimate, 0.025), 
    ci_upper = quantile(estimate, 0.975)) %>% 
  knitr::kable()
```

R-squared values 
```{r}
boot_rsqd <- weather_bootstraps %>%  
                        mutate(
                            models = map(strap_sample, ~lm(tmax ~ tmin, data = .x) ),
                            summary = map(models, broom::glance), 
                            results = map(models, broom::tidy)) %>% 
                        select(-strap_sample, -models, -results) %>% 
                        unnest(summary) %>% 
                        select(strap_number, r.squared, adj.r.squared)

head(boot_rsqd) %>% knitr::kable(digits = 3)
```

log(B0*B1) values 
```{r}
logb_df <- results %>% 
                  select(-std.error, -statistic, -p.value) %>% 
                  pivot_wider(
                    names_from = term, 
                    values_from = estimate) %>% 
                  rowwise(strap_number) %>% 
                  mutate(
                    B0B1 = `(Intercept)` * tmin, 
                    logB0B1 = log(B0B1)) %>% 
                  select(strap_number, logB0B1) 

head(logb_df) %>% knitr::kable(digits = 3)
```

Confidence intervals for r-squared and log(B0*B1)
```{r}
combined_df <- full_join(boot_rsqd, logb_df, by = "strap_number")

combined_df <- combined_df %>% 
                      pivot_longer(cols = !strap_number, 
                                   names_to = "measure", 
                                   values_to = "estimate") %>% 
                      group_by(measure) %>% 
                      summarize(
                        ci_lower = quantile(estimate, 0.025), 
                        ci_upper = quantile(estimate, 0.975))

head(combined_df) %>% knitr::kable(digits = 3)
```

## Graph of bootstrap samples
```{r}
weather_bootstraps %>% 
  unnest(strap_sample) %>% 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_line(aes(group = strap_number), stat = "smooth", method = "lm", se = FALSE, alpha = .1, color = "blue") +
  geom_point(data = weather_df, alpha = .5) + 
  ggpubr::stat_regline_equation(label.y = 50, aes(label = ..rr.label..))
```
The distribution of the boostrap estimates is wider at the low-end of the distribution (by tmin = -10). The distribution narrows towards the higher end of the distribution (around tmin = 10 - 18 adn tmax = 18 - 25). At the highest end, the distribution appears to start widening again. 
